Moment Constraints Make Learning Latent-Variable Models Easier

### 2014-01-29 doc/icml2014/background.tex
Updated directed section

Let $\sG$ be a discrete graphical model with variables $V$, of which
  $O \subseteq V$ be observed and $H = V \setminus O$ be hidden.
Let $\Pa(x) \subset V$ be the parents of a variable $x$ in a directed
  model, and $\sN(x) \subset V$ be the neighbours of $x$ in an undirected
  model.
We will use $o_1, o_2, \cdots \in O$ to denote observed variables and
  $h_1, h_2, \cdots \in H$ to denote hidden variables.
For ease of exposition, we assume that the domain of all the observed
  variables is $[d]$ and the domain of all the hidden variables is
  $[k]$.

Define $\otimes$ to be the tensor product.
To denote the moments of observed variables, we will use the notation
\begin{align*}
  \mO_i &\eqdef \Pr(x_i), &
  \mO_{ij} &\eqdef \Pr(x_i, x_j), &
  \mO_{ijk} &\eqdef \Pr(x_i, x_j, x_k),
\end{align*}
and so on.
We will view $\mO_i, \mO_{ij}, \mO_{ijk}, \cdots$ as tensors of
  order 1, 2, 3, \ldots in $\Re^D, \Re^{D\times D}, \Re^{D \times
  D \times D}$.
For the moments of hidden variables, we will use analogous notation
\begin{align*}
  \mH_i &\eqdef \Pr(h_i), &
  \mH_{ij} &\eqdef \Pr(h_i, h_j), &
  \mH_{ijk} &\eqdef \Pr(h_i, h_j, h_k).
\end{align*}

@@

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $\sD = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.

% Maximum likelihood
The standard estimation procedure is maximum (marginal) likelihood,
  \begin{align*}
    \sL &\eqdef \max_{\theta \in \Re^d} \E_{x} \log p_\theta(x) \\
        &=      \max_{\theta \in \Re^d} \E_{x} \log \sum_z p_\theta(z),
  \end{align*}
  which is statistically efficient but computationally intractable
  because we must marginalize over latent variables $z$.
In practice, one uses gradient-based optimization procedures (e.g., EM
  or L-BFGS) on the marginal likelihood, which can get stuck in local
  optima.

@@

Recently, \citet{anandkumar12moments} presented a consistent algorithm
  to learn the parameters of a three-view mixture model
  (\figureref{three-view}) that sidesteps the local optima problem
  altogether. 
In the model, we have three observed variables $x_1, x_2, x_3 \in [d]$
  that are independent conditioned on the latent variable $h \in [k]$. 
  The parameters of this model are the conditional means $O^{(i)} \eqdef
  \E[x_i | h] \in \Re^{D \times K}$ for $i \in \{1,2,3\}$ and the mixture probabilities $\pi
  \in \Delta_{k-1}$.
  The key idea is that the observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and
  $\mO_{123}$ admit a factorization in terms of parameters;
\begin{align*}
  \mO_{ij} &= \sum_h \pi_h O^{(i)}_h \otimes O^{(j)}_h & \forall i,j \in \{1,2,3\} \\
  \mO_{123} &= \sum_h \pi_h O^{(1)}_h \otimes O^{(2)}_h \otimes O^{(3)}_h.
\end{align*}

\citet{anandkumar13tensor} presented a robust tensor power method, which
  we call $\TensorFactorize$, which converts consistent estimates of the observed moments 
  into consistent estimates of the parameters,
  $\pi, \mO_1, \mO_2, \mO_3$ (up to permutation of the columns), provided
  the following assumption holds;

\begin{assumption}
\label{asm:full-rank}
Assume the conditional moments $\mO_1, \mO_2, \mO_3$ have full column rank
  $k$, and that $\pi \succ 0$.
\end{assumption}

@@

\todo{Mention {\em method of moments} somewhere?}

>>>>

Let $\sG$ be a discrete graphical model with variables $V$, 
  let the subset $X \subseteq V$ be observed and 
  let $H = V \setminus O$ be hidden.
Let $\Pa(v) \subset V$ be the parents of a variable $v$ in a directed
  model, and $\sN(v) \subset V$ be the neighbours of $v$ in an undirected
  model.
We will use $x_1, x_2, \cdots, x_N \in X$ to denote observed variables and
  $h_1, h_2, \cdots, h_M \in H$ to denote hidden variables.
For ease of exposition, we assume that $x_i \in \{1, \cdots, d\}$ for
  all $i$, and that $h_i \in \{1, \cdots, k\}$ for all i.

We assume that the model $\sG$ is parameterized at every clique $\sC$.
If $\sG$ is directed, the parameters correspond to
  the conditional probability tables for each clique.
If $\sG$ is undirected, the parameters correspond to the clique
  potentials.

@@

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $\sD
  = (\vx^{(1)}, \dots, \vx^{(n)})$ where each $\vx^{(i)} \sim p_{\theta^*}$ for
  some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ that
  approximates $\theta^*$.

% Maximum likelihood
The standard estimation procedure is maximum (marginal) likelihood,
  \begin{align*}
    \sL &\eqdef \max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \\
        &=      \max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} p_\theta(\vx,\vh)
  \end{align*}
  which is statistically efficient but computationally intractable
  because we must marginalize over latent variables $z$.
In practice, one uses gradient-based optimization procedures (e.g., EM
  or L-BFGS) on the marginal likelihood, which can get stuck in local
  optima.

\paragraph{Notation}

We use $\otimes$ to denote the tensor product, e.g if $u \in \Re^k,
  v \in \Re^d$, then $u \otimes v \in \Re^{k \times d}$.
For a $\ell$-th order tensor $T \in \Re^{d \times \ldots \times
  d}$ and vectors $x_1, \cdots, x_\ell \in \Re^{d},$, we define 
\begin{align*}
  T(x_1, \cdots, x_\ell) 
      &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
            T_{i_1 i_2 \ldots i_l} ({x_1})_{i_1} \cdots ({x_\ell})_{i_\ell}.
\end{align*}
Analogously, for a matrices $M_1 \in \Re^{d \times k}, \cdots,
  M_\ell \in \Re^{d \times k}$, we define
\begin{align*}
  T(M_1, \cdots, M_\ell)_{j_1, \ldots, j_l} 
      &= \sum_{i_1 = 1}^{d} \cdots \sum_{i_\ell = 1}^{d} 
            T_{i_1 i_2 \ldots i_l} ({M_1})_{i_1j_1} \cdots ({M_\ell})_{i_\ell j_\ell}.
\end{align*}

To denote the moments of observed variables, we will use the notation
\begin{align*}
  \mO_i &\eqdef \Pr(x_i), &
  \mO_{ij} &\eqdef \Pr(x_i, x_j), &
  \mO_{ijk} &\eqdef \Pr(x_i, x_j, x_k),
\end{align*}
and so on.
We will view $\mO_i, \mO_{ij}, \mO_{ijk}, \cdots$ as tensors of
  order 1, 2, 3, \ldots in $\Re^d, \Re^{d\times d}, \Re^{d \times
  d \times d},\ldots$.
For the moments of hidden variables, we will use 
\begin{align*}
  \mH_i &\eqdef \Pr(h_i), &
  \mH_{ij} &\eqdef \Pr(h_i, h_j), &
  \mH_{ijk} &\eqdef \Pr(h_i, h_j, h_k).
\end{align*}

@@

Recently, \citet{anandkumar12moments} presented a consistent algorithm
  using the method of moments to learn the parameters of a three-view
  mixture model (\figureref{three-view}), an important case of directed
  graphical models, that sidesteps the local optima problem altogether. 
In the model, we have three observed variables $x_1, x_2, x_3$
  that are independent conditioned on a latent variable $h$.  
The parameters of this model are the mixture probabilities $\pi \eqdef
  \Pr(h) \in \Delta_{k-1}$ and the conditional moments $O^{(i)} \eqdef
  \Pr(x_i \given h) \in \Re^{d \times k}$ for $i \in \{1,2,3\}$.

  The key idea is that the observed moments $\mO_{12}, \mO_{23}, \mO_{13}$ and
  $\mO_{123}$ admit a factorization in terms of parameters;
\begin{align*}
  \mO_{ij} &= \sum_h \pi_h {O\oft i}^T_h \otimes {O\oft j}^T_h & \forall i,j \in \{1,2,3\} \\
  \mO_{123} &= \sum_h \pi_h {O\oft 1}^T_h \otimes {O\oft 2}^T_h \otimes  {O\oft 3}^T_h.
\end{align*}

\citet{anandkumar13tensor} presented a robust tensor power method, which
  we call $\TensorFactorize$, which converts consistent estimates of the observed moments 
  into consistent estimates of the parameters, $\pi, O\oft{1},
  O\oft{2}, O\oft{3}$ (up to permutation of the columns), provided
  the following assumption holds:

\begin{assumption}
\label{asm:full-rank}
Assume the conditional moments $O\oft{1}, O\oft{2}, O\oft{3}$ have full column rank
  $k$, and that $\pi \succ 0$.
\end{assumption}

@@

### 2014-01-29 doc/icml2014/directed.tex
Updated directed section

We now show that once we recover the observation
  potentials, identifying other clique potentials in a directed graphical
  model can be consistently estimated in a systematic manner.
In section \sectionref{undirected} we generalize our results to
  undirected log-linear models.

@@

Let us illustrate our approach with an example. 
Consider the directed grid model shown in \figureref{grid} which
  captures the different dependency structures possible in a Bayesian network.
The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  hidden variables $h_1, \cdots, h_4$.
The parameters of this model are the conditional probability tables
  $\pi, T, V$ and $O$; for the sake of simplicity, we assume that all
  the observation potentials are the same, i.e. $O \eqdef \Pr(x^\cdot_i
  | h_i)$ for all $i$, and that $\Pr(h_2 | h_1) = \Pr(h_3 | h_1) = T$.
Finally, we assume $O$ and $T$ have full column rank.

\paragraph{Estimating $O$}

Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$, allowing us to use
  \TensorFactorize as reviewed in \sectionref{setup}, to recover
  the moments $O$.
We can aggregate statistics from additional sets of independent observed
  variables, e.g. $\{x^a_2, x^b_2, x^a_3\}$, to arrive at a better
  estimate of $O$.

@@

\paragraph{Estimating $\pi$}

We can relate the state of $h_1$ to the moments of $x^a_1$; $\mO_1
  \eqdef \Pr(x^a_1) = O \pi$. From \assumptionref{full-rank}, $O$ has full
  column rank and thus can be inverted to recover $\pi$; $\pi = \pinv O
  \mO_1$.

\paragraph{Estimating $T$}

Similarly, the we can write down the moments of $x^a_1, x^a_2$ in terms
of the pairwise marginals $\mH_{12} \eqdef \Pr(h_1, h_2)$ and solve for
$\mH_{12}$;
\begin{align*}
  \mO_{12} &= O \mH_{12} O^T \\
  \mH_{12} &= \pinv O \mO_{12} (\pinv O)^T.
\end{align*}

@@

Finally, we can estimate $V$ from the third-order moments $\mO_{234}$ as
follows,
\begin{align*}
  \mO_{234} &= \mH_{234}(O, O, O) \\
  \mH_{234} &= \mO_{234}(\pinv O, \pinv O, \pinv O).
\end{align*}

@@

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnFactors}
  \label{algo:directed}
  \begin{algorithmic}
    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
    \ENSURE Parameters $\theta$ for $\sG$

      \FOR{$h \in H$} 
        \STATE Apply \TensorFactorize for each bottleneck $h$ and learn parameters $\Pr(x \mid h) ~ \forall x \in \sB(h)$.

%    \COMMENT{Recover observation potentials $O$ using bottlenecks}
      \ENDFOR
%      \COMMENT{\textbf{Step 2:} Recover clique potentials from the piecewise likelihood.}
      \FOR{every clique $\sC = \{h_1, \cdots, h_m\} \in \sG$} 
      \STATE Let $x_\sC = \{x_1, \cdots, x_m\}$.
      \STATE $\mH_\sC = \mO_{x_\sC}( \pinv O_{x_1}, \cdots, \pinv O_{x_m} )$.
%      Run expectation-maximization to convergence on the piecewise likelihood \eqref{eqn:piecewise}, over data $\{\vec x_\sC : x \in \sD\}$
      \ENDFOR
  \end{algorithmic}
\end{algorithm}

\algorithmref{directed} presents the natural extension of this approach.
Let us now describe the general properties for which the algorithm,
  $\LearnFactors$ works.

For the first step of the algorithm, we wish to identify conditional
  moments between every hidden variable and atleast one observed
  variable.
We call hidden variables with this property ``bottlenecks'',
  and define them as follows.
\begin{definition}(Bottleneck)
  A hidden variable $h$ is said to be a bottleneck if there exists at
    least three conditionally independent observed variables $o_1, o_2,
    o_3$. 
  Let $\vec x_\sB(h)$ be the set of conditionally independent observed
    variables for $h$.
\end{definition}

We will further assume that each observed variable has a single parent. 
This leads to the following two properties.
\begin{property}
  \label{prop:unique-parent}
  Every observed variable $o$ has exactly one parent, $|\Pa(o)| = 1$.
\end{property}
\begin{property}
  \label{prop:bottleneck}
  Every hidden variable is a bottleneck.
\end{property}

We note that \propertyref{unique-parent} can be relaxed if another
  algorithm is used to recover the conditional moments. For example,
  \citet{halpern13noisyor} propose an algorithm that can recover the
  conditional moments for a broad, but less general class of noisy-or
  networks.
\propertyref{bottleneck} can also be relaxed if factors share the same
  parameters. 
For example, because we can identify the observation potential from just
  $x^a_1, x^b_1$ and $x^a_2$, we do not require that $h_2, h_3$ or $h_4$
  be bottlenecks, and can correspondingly omit the observations $x^b_2,
  x^b_3$ and $x^b_4$.

Given \propertyref{unique-parent} and \propertyref{bottleneck}, step
  1 of \algorithmref{directed} is guaranteed to recover potentials
  $O^{(h)}_{x}$ for every hidden variable $h$ and observed variable $x \in
  \vec x_\sB(h)$.
\propertyref{bottleneck} guarantees that for every clique $\sC
  = \{h_1, \cdots, h_m\}$ in the graphical model $\sG$, there exists at
  least one bottleneck observable that depends on exactly one $h_i \in
  \sC$; let $x_{h_i}$ be that observed variable\verify. 
Finally, define $\vec x_\sC = \{x_{h_1}, \cdots x_{h_m}\}$ to be the
  participating observed variables in the clique $\sC$.

Finally, we have the following moment condition,
\begin{align*}
  \mO_{x_\sC} &= \mH_{\sC}(O_{x_{h_1}}, \cdots, O_{x_{h_m}}) \\
  \mH_{\sC} &= \mO_{x_\sC}(\pinv O_{x_{h_1}}, \cdots, \pinv O_{x_{h_m}}).
\end{align*}

This algorithm is simple and intuitive. Next, we characterize its sample complexity.

\begin{lemma}(Sample complexity for $\sC$)
We require some number of samples to learn the factor for $\sC$.
\end{lemma}

This gives us our final result for a graphical model in our model
family.
\begin{theorem}(Learnability of Bottleneck Graphs)
The parameters for a directed graphical $\sG$ with tree-width $k$
satisfying \propertyref{unique-parent}, \propertyref{bottleneck} and
\assumptionref{full-rank} can be learned with $O(\cdot)$ samples.
\end{theorem}

\paragraph{Aggregating observations}

When we have multiple observed variables for a hidden variable, it is
  always beneficial to aggregate over them.
Note that $O_h$ will have rank atleast that of $O$; this encodes the
  intuitive fact that we might be able to identify $h$ from multiple
  observed variables even if we can not identify it from a single
  observed variable. 
Suppose $x_\sB(h) = \{x_1, x_2, \cdots, x_l\}$, let $O_h \in
  \Re^{D^l \times K}$ be an
  aggregated observation potential, such that $O_h(x_1, x_2, \cdots,
  x_l) = \mH(x_1|h) \mH(x_2|h) \cdots \mH(x_l|h)$.
Given \assumptionref{full-rank}, $O_h$ have rank $K$.
After step 1, $O_h$ is a known quantity.

>>>>

Once we recover the conditional moments $O\oft{i,j} \eqdef \Pr(x_i
  | h_j)$ for some $i,j$, we present a systematic approach to learn the
  conditional probability tables for every clique for a class of directed
  graphical models.
We begin by illustrating our approach with an example, followed by
  describing the simple conditions for which the approach works and some
  theoretical properties.

@@

Consider the directed grid model shown in \figureref{grid} which
  captures all the different dependency structures possible in
  a Bayesian network and is hence a good example for intuition\reword.
The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  hidden variables $h_1, \cdots, h_4$.
The parameters of this model are the conditional probability tables
  $\pi \eqdef \Pr(h_1), T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1),
  V \eqdef \Pr(h_4 | h_2, h_3)$ and $O \eqdef \Pr(x^a_i | h_i)
  =  \Pr(x^b_i | h_i)$. 
Finally, we assume $O$ and $T$ have full column rank.

\paragraph{Estimating $~O$}

Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$, allowing us to use
  \TensorFactorize, as reviewed in \sectionref{setup}, to recover
  the $O$.
We can exploit parameter sharing between the different $x$ to aggregate
  statistics from additional sets of independent
  observed variables, or {\em views}, e.g. $\{x^a_2, x^b_2, x^a_3\}$, to
  arrive at a better estimate of $O$.

@@

\paragraph{Estimating $~\pi$}

The moments of $x^a_1$,$\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
  $\pi$; $\mO_1 = O \pi$. 
By \assumptionref{full-rank}, $O$ has full column rank and thus can be
  inverted to recover $\pi$; $\pi = \pinv O \mO_1$.

\paragraph{Estimating $~T$}

Similarly, we can write down the moments of $x^a_1, x^a_2$, $\mO_{12}
  \eqdef \Pr(x^a_1, x^a_2)$ in terms of the pairwise marginals $\mH_{12}
  \eqdef \Pr(h_1, h_2)$ and solve for $\mH_{12}$;
\begin{align*}
  \mO_{12} &= O \mH_{12} O^T \\
  \mH_{12} &= \pinv O \mO_{12} \pinvt O.
\end{align*}

@@

Finally, we can estimate $V$ from the third-order moments $\mO_{234} \eqdef \Pr(x^a_2, x^a_3, x^a_4)$ as
follows,
\begin{align*}
  \mO_{234} &= \mH_{234}(O, O, O) \\
  \mH_{234} &= \mO_{234}(\pinv O, \pinv O, \pinv O).
\end{align*}

@@

Fundamentally, our algorithm takes advantage of having a view for every
  hidden variable in a clique to help identify its hidden state. The
  ideal property we wish our graphical models to possess is the
  following:
\begin{property}(Exclusive views)
  \label{prop:exclusive-views}
For every clique $\sC$ of $\sG$ and every hidden variable $h \in \sC$,
  there is some observed variable, $x_{v_h}$, which is conditionally
  independent of the rest of the clique given $h$ and whose conditional
  moment $\mOpp{v_h}{h} \eqdef \Pr(x_{v_h} | h)$ is known. 
We call $x_{v_h}$ the exclusive view of $h$ in $\sC$.
\end{property}

Before we identify a family of models which entail this property, let us
  show that it is sufficient to learn the parameters of $\sG$.

Consider any clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$. Let $x_{v_j}$ be the exclusive
  view for $h_{i_j}$ in $\sC$ and $\sV = \{x_{v_1}, \cdots, x_{v_m}\}$
  be the set of views for the clique $\sC$.
We can write down the moments $\mO_\sV$,
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{h_{i_1} = 1}^{d} \cdots \sum_{h_{i_1} = 1}^{d} 
      \Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
      &= \sum_{h_{i_1} = 1}^{d} \cdots \sum_{h_{i_1} = 1}^{d} \Pr(\Sh{\sC}) 
          \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
\end{align*}

If each $\mOpp{v_j}{i_j}$ has full column rank, then we can recover the
hidden marginals $Z_\sC$ by inverting the conditional moments,
\begin{align*}
  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}).
\end{align*}
Finally, the conditional probability tables for $\sC$ can easily be gotten from
  $Z_\sC$ via conditioning.

For each $\mOpp{v_j}{i_j}$ to have full column rank, it is equivalent
  to the following generalized assumption of
  \assumptionref{full-rank},

\begin{assumption} 
  \label{asm:full-rank-plus}
  For every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$, the
  marginal distribution $Z_\sC$ derived by the parameters is full rank
  for every mode-unfolding of $Z_\sC$.
\end{assumption}

Intuitively, this assumption says that the marginal probability of every state for any hidden variable is non-zero.

\begin{lemma}
  \label{lem:goodness}

  Given \assumptionref{full-rank-plus}, (a) the marginal distribution
  $\Pr(h_i) \succ 0$ for every $h_i$ in the graph and (b)
  $\mOpp{v_j}{i_j}$ has full column rank.
\end{lemma}
\begin{proof}
  By variable elimination.
\end{proof}

\paragraph{Required structural properties}

We will now provide a sufficient condition for
\propertyref{exclusive-views} to hold. 
Recall that given three observed
variables, $x_1, x_2, x_3$ that are independent conditioned on $h$,
\TensorFactorize allows us to recover the conditional moments $\mOp{1},
\mOp{2}, \mOp{3}$. We define a hidden variable that has this capability to be a bottleneck.
\begin{definition}(Bottleneck)
  A hidden variable $h$ is said to be a bottleneck if there exists at
    least three conditionally independent observed variables, or views,
    $x_1, x_2, x_3$. 
  Let $\xB{h}$ be the set of views for $h$.
\end{definition}

The only property a graphical models must have is the following,
\begin{property}
  \label{prop:bottleneck}
  Every hidden variable is a bottleneck.
\end{property}

We note that \propertyref{bottleneck} can be relaxed if some cliques
  share parameters;
For example, because we can identify the observation potential from just
  $x^a_1, x^b_1$ and $x^a_2$, we do not require that $h_2, h_3$ or $h_4$
  be bottlenecks, and can correspondingly omit the observations $x^b_2,
  x^b_3$ and $x^b_4$.
In this case, we only require that the hidden variables in distinct
  cliques share parameters.

It is not immediately obvious that \propertyref{bottleneck} implies
  \propertyref{exclusive-views}. 
Another consideration is that while
  \TensorFactorize gives us the conditional moments for every
  bottleneck, it can not recover the conditional probability table when
  there are multiple parents for an observed variable.
The following lemma shows us how to handle this case.

\begin{lemma}(Canonical form) 
  Every directed graphical model can be transformed into one in which
  the observed variables are leaves with exactly one parent.
\end{lemma}
\begin{proof}
  \begin{figure}
    \label{fig:reduction}
    \centering
    \subimport{figures/}{reduction.tikz}
    \caption{Reduction to canonical form.}
  \end{figure}

  Let $x$ be an observed variable with parents $h_1, \cdots, h_p$ and
  children $h_{p+1}, \cdots, h_c$. Replace $x$ with a phantom hidden
  variable $h_x$, and make $x$ the child of $h_x$ (\figureref{reduction}). The domain of $h_x$
  has the same as $x$, and $\mOpp{x}{h_x} = I$.

  The parameters for $h_x$ are the same as $x$.
\end{proof}

Applying \TensorFactorize to every bottleneck gives us candidate views
  for every hidden variable. Now we show that every clique also has
  exclusive views.

\begin{lemma}(Bottlenecks guarantee exclusive views)
  Provided \propertyref{bottleneck} holds, every hidden variable
    $h_{i_1}, \cdots, h_{i_m}$ in a clique $\sC$ has a exclusive view.
\end{lemma}
\begin{proof}
  Let $h$ be a hidden variable in $\sC$, let $\sV_{h}$ be the set of
    views for it and let $\sC^- \eqdef \sC \setminus \{h\}$.
  By definition, the views of $h$ are conditionally independent, viz.
    every d-connecting path between two views $x_1, x_2 \in \sV_h$ must pass through
    $h$.
  A view $x \in \sV_{h}$ is exclusive for $h$ iff $x \not\in \sV_{h'} ~ \forall h'
  \in \sC^-$

  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in \sC^-} \sV_{h'}$. 
  However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
    exists some $h_1, h_2 \in \sC^-$ (not necessarily distinct) such that
    $x_1$ is d-connected to $h_1$ and $x_2$ is d-connected $h_2$.
  By participation in the clique, either $h_1 = h_2$ or $h_1$ is
    d-connected to $h_2$, implying that $x_1$ and $x_2$ are d-connected
    via a path that does not pass through $h$, contradicting the
    statement that $x_1, x_2 \in \sV$.

  Thus it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
    h \in \sC} \sV_{h'}$ and hence there exists a unique view for $h$.

  Algorithmically, finding the unique views can be done by subtracting
    out the remaining set, i.e. $x_{v_h} \in \sV_h \setminus \Union_{h'
    \in \sC} \sV_{h'}$
\end{proof}

\paragraph{Aligning the views}
\todo{How do we make sure all the recovered $O$ are align? Shouldn't be Hard}

With this lemma in place, we present the full algorithm, \LearnMarginals
in \algorithmref{directed}.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnMarginals}
  \label{algo:directed}
  \begin{algorithmic}
    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
    \ENSURE Marginals $Z_\sC$ for every clique $\sC \in \sG$

      \FOR{$h_i \in H$} 
        \STATE Apply \TensorFactorize to learn conditional moments
        $\mOpp{v}{i}$ for every $x_v \in \xB h$.

%    \COMMENT{Recover observation potentials $O$ using bottlenecks}
      \ENDFOR
%      \COMMENT{\textbf{Step 2:} Recover clique potentials from the piecewise likelihood.}
\FOR{every clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$} 
\STATE Identify unique views $\Sx{\sV} = \{x_{v_1}, \cdots, x_{v_m}\}$.
\STATE $\mH_\sC \gets \mO_{\Sx{\sV}}( \pinv{\mOpp{v_1}{i_i}}, \cdots, \pinv{\mOpp{v_m}{i_m}} )$.
      \ENDFOR
  \end{algorithmic}
\end{algorithm}

\paragraph{Theoretical properties}

As each step of \LearnMarginals is a consistent estimator, it is
  straightforward that as whole it is a consistent estimator for $\sG$.
We describe its asymptotic complexity to see what the dependence on the
  parameters of the problem are.

\begin{lemma}(Sample complexity for $\sC$)

We require some number of samples to learn the factor for $\sC$.
\end{lemma}
\begin{proof}

For the sake of simplicity, let us assume that our dataset of $n$
  samples is partitioned into two sets.
  The first $\bytwo{n}$ samples are used to estimate $\mOpphat{v}{i}$ for
  every $h_i \in H$ and $x_v \in \sV_{h}$ in step 1 of \algorithmref{directed}.
\citet{anandkumar13tensor} show that 
\begin{align*}
  \|\mOpphat{v}{i} - \mOpp{v}{i}\|_F &= O( \frac{k \pi_{\max}^2}{\sigma_k(M_2)^5} ). 
\end{align*}

The next $\bytwo{n}$ samples are used to compute the moments $M_\sV$ for
  every clique in step 2.

We have for each clique,
\begin{align*}
  Z_{\sC} &= \mO_\sV(\pinv{\mOpp{v_1}{i_1}},\cdots,\pinv{\mOpp{v_m}{i_m}}) \\
\end{align*}

Choosing to represent $Z_\sC$ as a vector and $\mOpp{\cdot}{\cdot}$ as a matrix,
let $\pinv{\mOpp{\sV}{\sC}} \eqdef \pinv{\mOpp{v_1}{i_1}} \otimes \cdots \otimes \pinv{\mOpp{v_m}{i_m}}$,
\begin{align*}
  Z_{\sC} &= \pinv{\mOpp{\sV}{\sC}} \mO_\sV.
\end{align*}

We know the asymptotic distribution of $M_\sV$ goes as,
\begin{align*}
  \sqrt{n}(\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV),
\end{align*}
where $\Sigma_\sV$ is the variance. Because $M_\sV$ describes the discrete marginal distribution, we have 
\begin{align*}
  \Sigma_\sV &= \diag  M_\sV  - M_\sV M_\sV^T,
\end{align*}

Thus, $Z_\sC$ goes as
\begin{align*}
  \sqrt{n}(\hat Z_\sV - Z_\sV) \convind \sN(0, \pinvt{\mOpp{\sV}{\sC}} \Sigma_\sV \pinv{\mOpp{\sV}{\sC}}).
\end{align*}

\todo{Convert from $Z_\sC$ to the actual CPTs}.

\end{proof}
